Performing Cluster Analysis on TAD boundaries & Annotations

### 2019-03-22 Scripts/evaluating_pca.Rmd
including grouping by kmedoid clustering in pca plots

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(factoextra)
library(FactoMineR)
library(ggfortify)
```

@@

```{r}
fileslst <- c(paste("Z:/TAD_data_analysis/GM12878/",c("10kb", "25kb", "50kb", "100kb"), "/Ensemble/PCA/pcadat_GM12878_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_count.rds",sep=""),
              paste("Z:/TAD_data_analysis/GM12878/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_GM12878_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_perc.rds",sep=""),
              paste("Z:/TAD_data_analysis/GM12878/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_GM12878_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_dist.rds",sep=""),
              paste("Z:/TAD_data_analysis/K562/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_K562_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_count.rds",sep=""),
              paste("Z:/TAD_data_analysis/K562/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_K562_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_perc.rds",sep=""),
              paste("Z:/TAD_data_analysis/K562/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_K562_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_dist.rds",sep=""))

#file names for heatmaps
heatnames <- c(paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_count", "_heatmap.png", sep=""),
               paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_perc", "_heatmap.png", sep=""),
               paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_dist", "_heatmap.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_count", "_heatmap.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_perc", "_heatmap.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_dist", "_heatmap.png", sep=""))

#file names for pca plots
scorepnames <- c(paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_count", "_scoreplots.png", sep=""),
               paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_perc", "_scoreplots.png", sep=""),
               paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_dist", "_scoreplots.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_count", "_scoreplots.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_perc", "_scoreplots.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_dist", "_scoreplots.png", sep=""))

##create annotation groups
cs <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/BroadHMM")))
hm <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")))

for(i in 1:length(fileslst)){
  print(i)
  pcadata <- readRDS(fileslst[i])

  #performing log2 transformation only for distance predictor types
  if(grepl("_dist", colnames(pcadata)[1])){
    pcadata <- apply(pcadata,2,function(x){log(x + 1, base=2)})
  }

  #filtering low variance predictors
  nzv <- nearZeroVar(pcadata, saveMetrics= TRUE)
  nzvar <- rownames(nzv[nzv$nzv,])
  if(length(nzvar) > 0){ 
    pcadata_f <- pcadata[, -which(colnames(pcadata) %in% nzvar)]
    }else{pcadata_f <- pcadata}

  #cleaning column names
  colnames(pcadata_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadata_f))

  #heatmap 
  my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
  png(filename = paste("C:/Users/stili/Documents/TAD_miscellaneous/figures/pca_plots/", heatnames[i], sep=""),
      width=11, height=8, units="in", res = 300)
  heatmap.2(pcadata_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadata_f),
          #rowsep=1:nrow(pcadata_f),
          keysize = 1) 
  dev.off()

  #PCA
  #pcadata_f <- t(pcadata_f)
  #Group <- ifelse(rownames(pcadata_f) %in% cs, "Chromatin State", ###chromatin states
  #                                          ifelse(rownames(pcadata_f)=="DNase", "DNase I", ###dnase
  #                                                 ifelse(rownames(pcadata_f) %in% hm, "Histone Modification", ###histone modifications
  #                                                        "TFBS"))) ###transcription factor binding sites
  #pcadata_f <- cbind.data.frame(pcadata_f, Group)
  #
  ###Plots without loadings
  #p <- autoplot(prcomp(pcadata_f[,-(dim(pcadata_f)[2])]),
  #       #loadings = TRUE,
  #       label = TRUE, 
  #       label.size = 3,
  #       label.repel=T,
  #       data=pcadata_f,
  #       colour='Group',
  #       label.show.legend = FALSE) +
  #theme_bw()

  #ggsave(p, filename = paste("C:/Users/stili/Documents/TAD_miscellaneous/figures/pca_plots/", 
  #                           scorepnames[i], sep=""),
  #       width=8, height=11, units="in", dpi=300)

@@

>>>>

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(factoextra)
library(FactoMineR)
library(ggfortify)
library(fpc)
```

@@

```{r}
pcaplotfunction <- function(cl, resolution, predictortype, annotation){

  #set up variable names to subset
  chromstatevars <- gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/BroadHMM"))
  histonevars <- gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/HistoneModifications"))
  if(cl=="GM12878"){
    tfbsvars <-  gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878"))
  } else {
    tfbsvars <-  gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/k562"))
  }

  if(cl=="GM12878"){
    datadir = paste("Z:/TAD_data_analysis/", cl, "/", resolution, "/Ensemble/PCA", 
                    "/pcadat_gm12878_", resolution, "_", predictortype, ".rds",
                    sep="")
  }else{datadir = paste("Z:/TAD_data_analysis/", cl, "/", resolution, "/Ensemble/PCA", 
                        "/pcadat_", cl, "_", resolution, "_", predictortype, ".rds",
                        sep="")}

  data <- readRDS(datadir)

  if(predictortype=="dist"){
    #performing log2 transformation on distances
    data <- apply(data,2,function(x){log(x + 1, base=2)})
  }

  colnames(data) <- gsub(paste("_", predictortype, sep=""),"",colnames(data))

  data <- t(data)

  if(annotation=="chromstates"){
    data <- data[which(rownames(data) %in% chromstatevars),];
    rownames(data) <- gsub(paste(c("Gm12878", "K562",
                                   "-",
                                   "_",
                                   "Haib",
                                   "Sydh",
                                   "Uta",
                                   "Uw",
                                   "Broad",
                                   "Uchicago"), collapse = "|"),"", rownames(data))
  } else if(annotation=="histones"){
    data <- data[which(rownames(data) %in% histonevars),];
    rownames(data) <- gsub(paste(c("Gm12878", "K562",
                                   "-",
                                   "_",
                                   "Haib",
                                   "Sydh",
                                   "Uta",
                                   "Uw",
                                   "Broad",
                                   "Uchicago"), collapse = "|"),"", rownames(data))
  } else if(annotation=="tfbs"){
    data <- data[which(rownames(data) %in% tfbsvars),];
    rownames(data) <- gsub(paste(c("Gm12878", "K562",
                                   "-",
                                   "_",
                                   "Haib",
                                   "Sydh",
                                   "Uta",
                                   "Uw",
                                   "Broad",
                                   "Uchicago"), collapse = "|"),"", rownames(data))
  }

  #determine optimal number of clusters
  pamk.best <- pamk(data)

  #perform PAM clustering
  pam.med <- pam(data, k = pamk.best$nc, diss = FALSE,
                 metric = "manhattan", 
                 #medoids = NULL, 
                 #stand = FALSE, 
                 cluster.only = FALSE,
                 do.swap = TRUE,
                 keep.diss = FALSE,
                 keep.data = FALSE,
                 pamonce = FALSE, 
                 trace.lev = 0)

  data <- data.frame(data)
  data$Cluster <- factor(pam.med$clustering)
  data$Medoid <- factor(ifelse(rownames(data) %in% rownames(pam.med$medoids), "Center", "Other"),
                        levels=c("Other", "Center"))

  png(filename = paste("C:/Users/stili/Documents/TAD_miscellaneous/figures/pca_plots/", 
                       cl, resolution, predictortype, annotation,
                       ".png", sep=""),
      width=11, height=8, units="in", res = 300)
  autoplot(prcomp(data[,-c(dim(data)[2]-1,dim(data)[2])]),
           data=data,
           #loadings = TRUE,
           label = TRUE, 
           label.size = 3,
           label.repel=T, 
           colour="Cluster",
           shape="Medoid",
           label.show.legend = FALSE) +
    theme_bw()
  dev.off()

@@

###GM12878

```{r}
pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "count", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "count", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "count", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "perc", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "perc", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "perc", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "dist", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "dist", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "10kb", predictortype = "dist", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "count", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "count", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "count", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "perc", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "perc", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "perc", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "dist", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "dist", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "25kb", predictortype = "dist", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "count", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "count", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "count", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "perc", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "perc", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "perc", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "dist", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "dist", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "50kb", predictortype = "dist", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "count", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "count", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "count", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "perc", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "perc", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "perc", annotation = "tfbs")

pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "dist", annotation = "chromstates")
pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "dist", annotation = "histones")
pcaplotfunction(cl = "GM12878", resolution = "100kb", predictortype = "dist", annotation = "tfbs")

###K562

```{r}
